### 2021-11-13 NSF2021/Description/broaderimpacts.tex

This project aims to advance the mathematical modeling of collective 
dynamics of amphiphilic particles. The simulations use a new, yet intuitive,
approach that can account for important and complex systems that are out of 
reach in computational material science. These complex systems include 
fusion and fission of amphiphilic bilayer membranes and optimal shape design
in metamaterials. The development of three-dimensional 
models describing colloidal systems could be transformative in biomedicine
and material science. The research draws from expertise in scientific 
computing, physics of fluids, and mathematics. The mathematical component 
incorporates leading techniques from geometric analysis and offers  
insight 
into fundamentals of material science. The project brings socially consequential research into the classroom, and offers undergraduates the opportunity to train alongside graduate students and postdoctoral fellows.  With its combination of mathematical modeling, analysis, and scientific computing, the project highlights the importance of mathematics and computation to all areas of science and engineering.
%The proposed mathematical analysis, modeling, and numerical algorithms
%will transform our understanding of the collective dynamics of
%amphiphilic particles such as (1) their self-assembly into micelles and
%bilayers, (2) the material properties of such self-assembly, and (3) the
%interaction between these building blocks. In addition to biophysical
%applications, amphiphilic Janus particles have recent popularity in the
%fabrication of smart materials. The proposed work will have a
%transformative impact on precision design for specific mechanical
%properties of materials made of amphiphilic nanoparticles. An important
%component of this proposal is the interdisciplinary education and
%training of both undergraduate students, graduate students, and
%postdoctoral fellows. The combination of mathematical modeling,
%analysis, and scientific computing in this project provides a compelling
%example of the importance of mathematics and computation in biophysics
%and engineering applications. The concepts and methods described here go
%beyond the context of amphiphilic particles and lipid molecules. They
%extend to other problems featuring microscopic phase separation that
%leads to formation of mesoscopic domains. This situation arises, for
%example, in biological development in systems biology. The methods
%developed here have the potential to impact those and other related
%areas in biomedicine and biotechnology.

>>>>

This project aims to advance the mathematical modeling of collective
dynamics of amphiphilic particles. The simulations use a new, yet
intuitive, approach that accounts for important and complex systems
that are out of reach in computational material science. These complex
systems include fusion and fission of amphiphilic bilayer membranes and
optimal shape design in metamaterials. The development of
three-dimensional models describing colloidal systems could be
transformative in biomedicine and material science. The research draws
from expertise in scientific computing, physics of fluids, and
mathematics. The mathematical component incorporates leading techniques
from geometric analysis and offers insight into fundamentals of material
science. The project brings socially consequential research into the
classroom, and offers undergraduates the opportunity to train alongside
graduate students and postdoctoral fellows. With its combination of
mathematical modeling, analysis, and scientific computing, the project
highlights the importance of mathematics and computation to all areas of
science and engineering.
%The proposed mathematical analysis, modeling, and numerical algorithms
%will transform our understanding of the collective dynamics of
%amphiphilic particles such as (1) their self-assembly into micelles and
%bilayers, (2) the material properties of such self-assembly, and (3) the
%interaction between these building blocks. In addition to biophysical
%applications, amphiphilic Janus particles have recent popularity in the
%fabrication of smart materials. The proposed work will have a
%transformative impact on precision design for specific mechanical
%properties of materials made of amphiphilic nanoparticles. An important
%component of this proposal is the interdisciplinary education and
%training of both undergraduate students, graduate students, and
%postdoctoral fellows. The combination of mathematical modeling,
%analysis, and scientific computing in this project provides a compelling
%example of the importance of mathematics and computation in biophysics
%and engineering applications. The concepts and methods described here go
%beyond the context of amphiphilic particles and lipid molecules. They
%extend to other problems featuring microscopic phase separation that
%leads to formation of mesoscopic domains. This situation arises, for
%example, in biological development in systems biology. The methods
%developed here have the potential to impact those and other related
%areas in biomedicine and biotechnology.

### 2021-11-13 NSF2021/Description/specificaim2.tex

\subsection{Specific Aim 2: Efficient, high-order methods for
large-scale simulations}
% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces and van der Waals
interactions to coordinate movement and bind microscopic components in
solvent~\cite{Pandey2011, Zhang2017, Siontorou2017}. Additionally,
hydrodynamic interactions are central to fabricate complex,
three-dimensional microstructures~\cite{Dasgupta2017, Leong2007,
Reynolds2019, Cho2010}. Hydrodynamic effects cannot be ignored since
they, like for the rates of biological functions like pore
dynamics~\cite{RYHAM20112929}, set the rate of dissipation. In Specific
Aim 2, we propose novel and efficient high-order numerical methods to
simulate dynamic assembly of particles under HAP potential in a viscous
solvent. As described in~\eqref{eq:stokes1}--\eqref{eq:stokes3} and
illustrated in Figure~\ref{fig:flow_map}, the solvent phase is modeled
by the Stokes equations for an incompressible, viscous fluid, and these
equations are coupled to the screened-Laplace equation~\eqref{eq:SL}
through viscous and hydrophobic stress balance.

%\label{subsec:specific_aim_2}
%\begin{wrapfigure}[14]{r}{0.61\textwidth}
%  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
%    \vspace{-15pt}
%    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
%    amphiphilic particles suspended in a solvent. The solvent velocity
%    satisfies the Stokes equations, and the hydrophobic forces depend on
%    the solvent activity that satisfies the screened Laplace equation.
%    The geometry is updated by solving the mobility problem.} 
%\end{wrapfigure}
The HAP model requires solving exterior problems for the order
parameter~\eqref{eq:SL} and the Stokes
equations~\eqref{eq:stokes1}--\eqref{eq:stokes3}. We will develop
methods based on combinations of volume and layer potentials. Numerical
methods to solve these integral equations (IEs) have several advantages
over their PDE-based counterpart: by construction, they satisfy
far-field and mass conservation conditions; only $\partial\Omega$ needs
to be discretized for layer potentials; carefully chosen IEs result in a
well-conditioned linear system that is solved with a mesh-independent
number of GMRES iterations; and high-order or spectral accuracy is
attainable with appropriate quadrature methods. The PIs have
demonstrated that IEs are a powerful tool to simulate two-dimensional
suspensions of amphiphilic particles~\cite{Fu2018_SIAM, FuQuRyYo20}.
Moreover, elements of these formulations have been described in three
dimensions~\cite{ying_2006, manasthesis, rac-gre2016}. This proposal
will extend the two-dimensional methods to more general amphiphilic
particle models that satisfy~\eqref{eq:SL}, and all the methods can
naturally be extended to three dimensions. \S\ref{subsec:AC} describes
how the non-linear Allen-Cahn equation~\eqref{eq:SL} will be iteratively
solved using IE formulations. Then, \S\ref{subsec:NumericalIssues}
describes how we will combine preconditioning strategies with fast
summation methods to accelerate the time to solution, and
\S\ref{subsec:timeStepping} describes adaptive time stepping methods,

>>>>

\subsection{Specific Aim 2: Efficient, high-order methods for
large-scale simulations}
% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces and van der Waals
interactions to coordinate movement and bind microscopic components in
solvent~\cite{Pandey2011, Zhang2017, Siontorou2017}. Additionally,
hydrodynamic interactions are central to fabricate complex,
three-dimensional microstructures~\cite{Dasgupta2017, Leong2007,
Reynolds2019, Cho2010}. Hydrodynamic effects cannot be ignored since
they, like for the rates of biological functions like pore
dynamics~\cite{RYHAM20112929}, set the rate of dissipation. In Specific
Aim 2, we propose novel and efficient high-order numerical methods to
simulate dynamic assembly of particles under HAP potential in a viscous
solvent. As described in~\eqref{eq:stokes} and illustrated in
Figure~\ref{fig:flow_map}, the solvent phase is modeled by the Stokes
equations for an incompressible, viscous fluid, and these equations are
coupled to the screened-Laplace equation~\eqref{eq:SL} through viscous
and hydrophobic stress balance.

%\label{subsec:specific_aim_2}
%\begin{wrapfigure}[14]{r}{0.61\textwidth}
%  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
%    \vspace{-15pt}
%    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
%    amphiphilic particles suspended in a solvent. The solvent velocity
%    satisfies the Stokes equations, and the hydrophobic forces depend on
%    the solvent activity that satisfies the screened Laplace equation.
%    The geometry is updated by solving the mobility problem.} 
%\end{wrapfigure}
The HAP model requires solving exterior problems for the order
parameter~\eqref{eq:SL} and the Stokes equations~\eqref{eq:stokes}. We
will develop methods based on combinations of volume and layer
potentials. Numerical methods to solve these integral equations (IEs)
have several advantages over their PDE-based counterpart: by
construction, they satisfy far-field and mass conservation conditions;
only $\partial\Omega$ needs to be discretized for layer potentials;
carefully chosen IEs result in a well-conditioned linear system that is
solved with a mesh-independent number of GMRES iterations; and
high-order or spectral accuracy is attainable with appropriate
quadrature methods. The PIs have demonstrated that IEs are a powerful
tool to simulate two-dimensional suspensions of amphiphilic
particles~\cite{Fu2018_SIAM, FuQuRyYo20}. Moreover, elements of these
formulations have been described in three dimensions~\cite{ying_2006,
manasthesis, rac-gre2016}. This proposal will extend the two-dimensional
methods to more general amphiphilic particle models that
satisfy~\eqref{eq:SL}, and all the methods can naturally be extended to
three dimensions. \S\ref{subsec:AC} describes how the non-linear
Allen-Cahn equation~\eqref{eq:SL} will be iteratively solved using IE
formulations. Then, \S\ref{subsec:NumericalIssues} describes how we will
combine preconditioning strategies with fast summation methods to
accelerate the time to solution, and \S\ref{subsec:timeStepping}
describes adaptive time stepping methods,
